---
title: "rmdDayOne"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(eval = FALSE, echo = FALSE)
```

# Funktionen 

## Gute Funktionen

- Einfach
- Kurz
- Abgeschlossen (self-contained)

## Zielsetzung

- Abstraktion
- Vermeidung von Redundanz

```{r eval = FALSE}
someFunction <- function(<args>) {
  <body>
} 
```

```{r}
someFunction <- function(x) {
  x
}
someFunction(1)
```

## Arguments

```{r}
args(someFunction)
formals(someFunction)
```


## Body

```{r}
body(someFunction)
```

## Environment

```{r}
environment(someFunction)
```


## Function arguments

- exact
- partial
- positional

```{r}
mydata <- c(rnorm(100), NA)

sd(mydata, na.rm = TRUE)
sd(na.rm = TRUE, mydata)
```

## Defaults

```{r}
someFunction <- function(x, mul = x) x * mul
someFunction(3)

tail <- function(df, selection = nrow(df)) df[selection, ] 
df <- data.frame(x = rnorm(10))
tail(df)
tail(df, 1:3)
```

## Missing Arguments

```{r, error=TRUE}
someFunction <- function(a, b) {
  cat(a, "\n")
  cat(b)
}

someFunction("a")

someFunction <- function(a) {
  cat(a, "\n")
  cat(b)
}

b <- "b"
someFunction("a")

a <-"test1"
b <- "test2"

newFunction <- function(a, b) {
  cat(a, "\n")
  cat(b)
}

```

## Ellipses Argument

```{r}
args(cat)
myplot <- function(x, y, type = "l", ...) plot(x, y, type = type, ...)
myplot(1:10, rnorm(10))
```



## Infix Functions

```{r}
"+" <- function(lhs, rhs) lhs - rhs
1 + 2
rm("+")

"%paste%" <- function(lhs, rhs) paste(lhs, rhs)
"left" %paste% 
  "right" %paste%
  "more"

"%paste%"("left", "right")
```

## Replacement Functions

```{r}
names(df)
names(df) <- "y"

df <- "names<-"(df, "y")
```


# Scoping Rules

- Welche Namen / Objekte können gefunden werden?
- Wann können diese Objekte gefunden werden?

```{r}

b <- function(x) {
  
  a <- function(y) {
    y^2
  }
  a(x)
}

b(2)
```


```{r}
y <- 2

b <- function(x, sep = "") {
  lapply(x, function(x) {
    sep
  })
}

b(10)
```


## Dynamic Lookup

```{r}
someFunction <- function(x) {
  someOtherFunction(x)
}

someOtherFunction <- function(y) y
someFunction(1)
```


## Name Masking

```{r}
nrow(df)
dim <- function(df) 2
nrow(df)

mynrow <- function(x) dim(x)[1L]
mynrow(df)
```


# Control Structures

  - **if, else:** testing a condition
  - **for:** execute a loop a fixed number of times
  - **while:** execute a loop while a condition is true
  - **repeat:** execute an infinite loop
  - **break:** break the execution of a loop
  - **next:** break the execution of a loop


## if, else

```{r, eval = FALSE}

if (condition) {
  ## do something
} else {
  ## do something else
}

```

```{r, error=TRUE}

sqrtOwn <- function(x){
  
  # does the input correspond with expectations
  if (x < 0) {
    stop("'sqrtOwn' can't handle negative values")
    
  } else {
    sqrt(x)
  }
}
  
sqrtOwn(16)
sqrtOwn(-4)

sqrtOwn(-3:1)

-3:1 < 0
```


```{r, eval = FALSE}
# or 
if (condition1) {
  ## do something
} else if (condition2) {
  ## do something different
} else {
  ## do something even more different
}
```


## for

```{r}
for(i in 1:5){
  
  print(i)
  
}
```

```{r}
x <- c("a", "b", "c", "d", "e")

for(i in 1:length(x)){  # index vector of corresponding length
  print(x[i])
}


for(element in x){ # element directly taken from vector
  print(element)
}

```


```{r}
data(iris)

for(column in iris){
  str(column)
}
```


### nested for-loops

```{r}
matA <- matrix(1:6, nrow = 2)

for(row in 1:nrow(matA)){
  for(col in 1:ncol(matA)){
    print(matA[row, col])
  }
}

```


### loops are suitable for iterative programming
```{r}
loopObject <- NULL

for(i in 1:10){
  loopObject <- c(loopObject, i^2)
}

```


## while 

```{r, eval = FALSE}
number <- 0

while(condition is true){
  # do something
}

```


```{r}
pos <- 0
step <- -1

while (pos >= -3 && pos <= 5) {
  
  cat("Step =", step <- step + 1, "\n")
  cat("Position =", pos, "\n")
  coin <- rbinom(n = 1, size = 1, prob = 0.5)
  if (coin == 1) { ## random walk
    pos <- pos + 1
  } else {
    pos <- pos - 1
  }
  
  if(step > 10) {
    warning("maximum amount of iterations")
    break
  }
  
}
```

```{r}
x <- "a"

is.numeric(x) && sqrt(x) > 2
is.character(x) || sqrt(x) > 2

```

## repeat

```{r, eval = FALSE}

repeat {
  
  # do something within every looping
  
  # quit loop, if
  if(condition1) break
  
}

```

# *apply Functions

## apply

```{r}
matB <- matrix(c(rnorm(199), NA), nrow = 20, ncol = 10)

apply(matB, MARGIN = 2, mean, na.rm = TRUE)
apply(matB, MARGIN = 1, sd)

apply(matB, MARGIN = 2, FUN = function(x) x[1])
matB[1, ]

```

works also for `data.frames`

```{r}
apply(iris[-5], 1, mean)
```


```{r}
# rowMeans, colMeans
# rowSums, colSums

colMeans(iris[-5])


```


## lapply

```{r, error = TRUE}
aList <- list(a = c(1:5, NA),
              b = rnorm(10),
              c = rpois(15,lambda = 3))

mean(aList$a)
mean(aList$b)

library(parallel)
mclapply(aList, mean, na.rm = TRUE, mc.cores = length(aList))

# alternative: for-loop
resList <- list()
for(elementName in names(aList)){
  resList[[elementName]] <-mean(aList[[elementName]], 
                                na.rm = TRUE)
}

```



```{r}
nElements <- 1:4

lapply(nElements, runif, n = 5, min = 0)

lapply(nElements, function(x) runif(n = 5, min = 0, max = x))


```

## sapply

```{r}
library(dplyr)
names(nElements) <- letters[1:4]

set.seed(1)
sapply(nElements, 
         function(x) runif(n = 5, min = 0, max = x),
         simplify = FALSE)



```

```{r}
charElements <- letters[1:5]
lapply(charElements, function(x) rnorm(1))

sapply(charElements, 
       function(x) rnorm(1))

# one has more control, if the data structure is
# generated afterwards with additional commands
resList <- sapply(charElements, 
                  function(x) rnorm(1), 
                  simplify = FALSE)

do.call(cbind, resList) # bind column-wise
do.call(rbind, resList) # bind row-wise

# # excursus
# runif(n = 10, min = 0, max = 5)
# do.call(runif, list(n = 10, 
#                     min = 0, 
#                     max = 5))


```


## mapply


```{r}
grep("a", letters)
grep("b", letters)
sapply(letters, function(x) grep(x, letters), simplify = FALSE)
lapply(letters, function(x) grep(x, letters))
```


```{r}

makeNames <- function(vorname, nachname, ...) {
  paste(vorname, nachname, ...)
  }

makeNames("Steffen", "Wagner", sep ="")

mapply(makeNames,
       nachname = c("Lohmann", "Wagner", "Warnholz"),
       vorname = c("Dietmar", "Steffen", "Sebastian"),
       MoreArgs = list(sep = c("_")),
       SIMPLIFY = FALSE)

```

## split

```{r}
data(iris)

irisList <- split(iris, f = iris$Species)

lapply(irisList, summary)

lmList <- lapply(irisList, function(dat) lm(Sepal.Length ~ Sepal.Width, dat))


lapply(lmList, summary)

sapply(lmList, function(x) summary(x)$adj.r.squared)

```


fließtext `r 2+2`
